---
title: "Nintendo Switch Sales"
output: 
  flexdashboard::flex_dashboard:
    orientation: row
    vertical_layout: fill
---

<style type="text/css">

.navbar-inverse .navbar-nav>.active>a, .navbar-inverse .navbar-nav>.active>a:hover, .navbar-inverse .navbar-nav>.active>a:focus {
color: #3b67f5;
background-color: #303030;
border-color: #002a4c;
font-weight: 900;
}

.chart-title {
    border-bottom: 1px solid #d7d7d7;
    color: #3b67f5;
    font-size: 20px;
    font-weight: 700;
    padding: 7px 10px 4px;
}

.heading-level1 {  /* chart_title  */
   color: #00467f;
    font-size: 18px;
    font-weight: 500;
}

h4, .h4 {
   color: #002a4c;
    font-size: 15px;
    font-weight: 550;
}

a:visited {
   color: rgb(50%, 0%, 50%);
}

</style>


```{r}
#import libraries
library(flexdashboard)
library(ggthemes)
library(ggrepel)
library(ggplot2)
library(scales)
library(readxl)
library(tidyr)
library(dplyr)
library(data.table)
library(stats)
library(xts)

###### DATA CLEANING + PREP ######
#Read Data
myData <- read_excel("NintendoSwitchSales.xls")
#Rename Columns
colnames(myData) <- c("Platform", 'Metric', '2017', '2018', '2019', '2020')
#Drop Metric Column
df = subset(myData, select = -c(Metric) )
#Multiply yearly stats by 10,000
df[] = lapply(df, FUN = function(x) if (is.numeric(x)) return(x * 10000) else return(x))
#Remove string 'Nintendo Switch ' from Platform column
df$Platform <- gsub("Nintendo Switch ","",as.character(df$Platform))
#Separate Platform Column into Platform and Region
df <- separate(data = df, col = Platform, into = c("Platform", "Region"), sep = "/")
#Convert data from wide to long
df <- gather(df, Year, Sales, '2017':'2020', factor_key=TRUE)

# Data Prep for first page
hardware <- subset(df, Platform == 'hardware')
software <- subset(df, Platform == 'software')

hardware <- hardware %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

software <- software %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

hardware$Cumulative <- cumsum(hardware$YearTotal)
software$Cumulative <- cumsum(software$YearTotal)

hardware <- hardware %>% mutate(Platform = 'hardware')
software <- software %>% mutate(Platform = 'software')

platformDf <- rbind(hardware, software) 

# Data prep for second page
region1 <- subset(df, Region == 'The Americas')
region2 <- subset(df, Region == 'Europe')
region3 <- subset(df, Region == 'Japan')
region4 <- subset(df, Region == 'Other')

region1 <- region1 %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

region2 <- region2 %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

region3 <- region3 %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

region4 <- region4 %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

# Create cumulative column
region1$Cumulative <- cumsum(region1$YearTotal)
region2$Cumulative <- cumsum(region2$YearTotal)
region3$Cumulative <- cumsum(region3$YearTotal)
region4$Cumulative <- cumsum(region4$YearTotal)

region1 <- region1 %>% mutate(Region = 'The Americas')
region2 <- region2 %>% mutate(Region = 'Europe')
region3 <- region3 %>% mutate(Region = 'Japan')
region4 <- region4 %>% mutate(Region = 'Other')

# need data.table format to create growth column
setDT(region1)
setDT(region2)
setDT(region3)
setDT(region4)

# Create growth column
region1[ , Growth := round((((YearTotal - shift(YearTotal)) / shift(YearTotal)) * 100), 1)]
region2[ , Growth := round((((YearTotal - shift(YearTotal)) / shift(YearTotal)) * 100), 1)]
region3[ , Growth := round((((YearTotal - shift(YearTotal)) / shift(YearTotal)) * 100), 1)]
region4[ , Growth := round((((YearTotal - shift(YearTotal)) / shift(YearTotal)) * 100), 1)]

# sets na's to zero
region1[is.na(region1)] <- 0
region2[is.na(region2)] <- 0
region3[is.na(region3)] <- 0
region4[is.na(region4)] <- 0

# combine 4 regions into new df
regionDf <- rbind(region1, region2, region3, region4)

totalSalesDf <- df %>% 
  group_by(Year) %>% 
  summarise(YearTotal = sum(Sales))

# Exponential Prediction Model using Holt Winters with beta and gamma off
tsdf <- ts(totalSalesDf$YearTotal, frequency=1, start=c(2017))
modell <- HoltWinters(tsdf, beta = FALSE, gamma = FALSE)
pred <- predict(modell, prediction.interval = TRUE)

ts_pred <- ts(c(modell$fitted[, 1], pred), start = 2017, frequency = 1)
testdf <- merge(as.xts(ts_pred), as.xts(tsdf))
names(testdf) <- c("predicted", "actual")

totalSalesDf$Year <- as.numeric(as.character(totalSalesDf$Year))
```
<!-- ######################################################################################################## -->
<!-- FIRST PAGE -->
<!-- ######################################################################################################## -->
Platform Sales
===

Row
---
### Overall Sales {.value-box}

```{r}
totalSales <- sum(df['Sales'])
totalSales <- paste(format(round(totalSales / 1e6, 1), trim = TRUE), "M")
valueBox(totalSales, caption = 'Total Sales (Units)', icon="fa-dollar-sign", color = '#5cbd71')
```

### Software Sales {.value-box}

```{r}
software <- subset(df, Platform == 'software')
softwareSales <- sum(software$Sales)
softwareSales <- paste(format(round(softwareSales / 1e6, 1), trim = TRUE), "M")
valueBox(softwareSales, caption = 'Total Software Sales (Units)', icon="fa-compact-disc", color = '#3643ff')
```

### Hardware Sales {.value-box}

```{r}
hardware <- subset(df, Platform == 'hardware')
hardwareSales <- sum(hardware$Sales)
hardwareSales <- paste(format(round(hardwareSales / 1e6, 1), trim = TRUE), "M")
valueBox(hardwareSales, caption = 'Total Hardware Sales (Units)', icon="fa-desktop", color = '#bd5c5c')
```


Row
---


### Cumulative Sales

```{r}
p <- ggplot(platformDf, aes(x = as.numeric(Year), y = Cumulative))
p + geom_area(aes(fill = Platform, color = Platform), alpha = 0.4) +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'Cumulative Platform Revenue', y = 'Total Sales (Units)', x = 'Fiscal Year' )
```


### Yearly Sales

```{r}
p <- ggplot(data = df)
p + geom_bar(stat = 'identity', aes(x = Year, y = Sales, fill = Platform), alpha = 0.7) +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  labs( title = 'Yearly Platform Revenue', y = 'Total Sales (Units)', x = 'Fiscal Year' )

```

<!-- ######################################################################################################## -->
<!-- SECOND PAGE -->
<!-- ######################################################################################################## -->
Regional Sales
===

Row
---

### Region1 {.value-box}

```{r}
r1max <- max(region1$Cumulative)
softwareSales <- paste(format(round(r1max / 1e6, 1), trim = TRUE), "M")
valueBox(softwareSales, caption = 'The Americas Sales (Units)', icon="fa-earth-americas", color = '#9959b5')
```

### Region2 {.value-box}

```{r}
r2max <- max(region2$Cumulative)
softwareSales <- paste(format(round(r2max / 1e6, 1), trim = TRUE), "M")
valueBox(softwareSales, caption = 'Europe Sales (Units)', icon="fa-earth-europe", color = '#3abdc2')
```

### Region3 {.value-box}

```{r}
r3max <- max(region3$Cumulative)
softwareSales <- paste(format(round(r3max / 1e6, 1), trim = TRUE), "M")
valueBox(softwareSales, caption = 'Japan Sales (Units)', icon="fa-earth-asia", color = '#7eb362')
```

### Region4 {.value-box}

```{r}
r4max <- max(region4$Cumulative)
softwareSales <- paste(format(round(r4max / 1e6, 1), trim = TRUE), "M")
valueBox(softwareSales, caption = 'Other Sales (Units)', icon="fa-earth-oceania", color = '#bd6a6a')
```

Row
---

### Cumulative Sales

```{r}
p <- ggplot(regionDf, aes(x = as.numeric(Year), y = Cumulative))
p + geom_area(aes(fill = factor(Region, levels=c("Other", "Japan", "Europe", "The Americas")), color = factor(Region, levels=c("Other", "Japan", "Europe", "The Americas"))), alpha = 0.4) +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'Cumulative Regional Revenue', y = 'Total Sales (Units)', x = 'Fiscal Year', fill = 'Region', colour = 'Region' )
```

### Yearly Sales

```{r}
p <- ggplot(data = df)
p + geom_bar(stat = 'identity', aes(x = Year, y = Sales, fill = factor(Region, levels=c("Other", "Japan", "Europe", "The Americas"))), alpha = 0.7) +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  labs( title = 'Yearly Regional Revenue', y = 'Total Sales (Units)', x = 'Fiscal Year' )

```


<!-- ######################################################################################################## -->
<!-- THIRD PAGE -->
<!-- ######################################################################################################## -->
Regional Growth
===

Row
---
### Growth in The Americas

```{r}
endPoint <- region1 %>% filter(Year == 2020)

p <- ggplot(region1, aes(x = as.numeric(Year), y = Growth))
p + geom_line(stat = 'identity') +
  geom_point(stat = 'identity') +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'The Americas\' Year over Year Growth', y = 'Growth (%)', x = 'Fiscal Year' ) +
  geom_text_repel(aes(label = Growth), data = endPoint, fontface ="plain", color = "black", size = 4, nudge_y = -2)
```

### Growth in Europe

```{r}
endPoint <- region2 %>% filter(Year == 2020)

p <- ggplot(region2, aes(x = as.numeric(Year), y = Growth))
p + geom_line(stat = 'identity') +
  geom_point(stat = 'identity') +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'Europe\'s Year over Year Growth', y = 'Growth (%)', x = 'Fiscal Year' ) +
  geom_text_repel(aes(label = Growth), data = endPoint, fontface ="plain", color = "black", size = 4, nudge_y = -2)
```

Row
---
### Growth in Japan

```{r}
endPoint <- region3 %>% filter(Year == 2020)

p <- ggplot(region3, aes(x = as.numeric(Year), y = Growth))
p + geom_line(stat = 'identity') +
  geom_point(stat = 'identity') +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'Japan\'s Year over Year Growth', y = 'Growth (%)', x = 'Fiscal Year' ) +
  geom_text_repel(aes(label = Growth), data = endPoint, fontface ="plain", color = "black", size = 4, nudge_y = -2)
```

### Growth in Other Regions

```{r}
endPoint <- region4 %>% filter(Year == 2020)

p <- ggplot(region4, aes(x = as.numeric(Year), y = Growth))
p + geom_line(stat = 'identity') +
  geom_point(stat = 'identity') +
  theme_grey() +
  theme(legend.position = 'none') +
  scale_x_continuous(labels = c('2017', '2018', '2019', '2020')) +
  labs( title = 'Other Regions\' Year over Year Growth', y = 'Growth (%)', x = 'Fiscal Year' ) +
  geom_text_repel(aes(label = Growth), data = endPoint, fontface ="plain", color = "black", size = 4, nudge_y = -2)
```

<!-- ######################################################################################################## -->
<!-- FOURTH PAGE -->
<!-- ######################################################################################################## -->
Predicted Sales
===

Row
---


### Linear 2021 Sales {.value-box}


```{r}
lmodel <- lm(YearTotal ~ Year, data = totalSalesDf)
l2021 <- as.numeric(predict(lmodel, data.frame(Year = 2021)))
l2021 <- paste(format(round(l2021 / 1e6, 1), trim = TRUE), "M")
valueBox(l2021, caption = 'Linear 2021 Sales (Units)', icon="fa-earth-oceania", color = '#5cbd71')
```

### Linear 2022 Sales {.value-box}


```{r}
lmodel <- lm(YearTotal ~ Year, data = totalSalesDf)
l2022 <- predict(lmodel, data.frame(Year = 2022))
l2022 <- paste(format(round(l2022 / 1e6, 1), trim = TRUE), "M")
valueBox(l2022, caption = 'Linear 2022 Sales (Units)', icon="fa-earth-oceania", color = '#5cbd71')
```

### Exponential 2021 Sales {.value-box}


```{r}
e2021 <- testdf[5,1]
e2021 <- paste(format(round(e2021 / 1e6, 1), trim = TRUE), "M")
valueBox(e2021, caption = 'Exponential 2021 Sales (Units)', icon="fa-earth-oceania", color = '#5cbd71')
```

### Exponential 2022 Sales {.value-box}


```{r}
e2022 <- testdf[6,1]
e2022 <- paste(format(round(e2022 / 1e6, 1), trim = TRUE), "M")
valueBox(e2022, caption = 'Exponential 2022 Sales (Units)', icon="fa-earth-oceania", color = '#bd6a6a')
```


Row
---


### Predicted Sales Linear


```{r}

p <- ggplot()
p + geom_line(totalSalesDf, mapping = aes(x = Year, y = YearTotal), stat = 'identity', size = 1.1) +
  geom_point(totalSalesDf, mapping = aes(x = Year, y = YearTotal), stat = 'identity', size = 1.1) +
  theme_grey() +
  geom_smooth(totalSalesDf, mapping = aes(x = Year, y = YearTotal), method = "lm", se = FALSE, fullrange=TRUE, alpha = 0.4, linetype = "dashed") +
  theme(legend.position = 'none') +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  labs( title = '2 Year Sales Linear Forecast', y = 'Total Sales (Units)', x = 'Fiscal Year') +
  geom_vline(xintercept=2020, linetype="dashed") +
  xlim(2017,2022)
```


### Predicted Sales Exponential


```{r}
ggplot(testdf, aes(x=as.POSIXct(index(testdf)))) +
  geom_line(aes(y=predicted), col='red') +
  geom_line(aes(y=actual), col='black') +
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6)) +
  theme_grey() +
  geom_vline(xintercept=as.numeric(as.POSIXct(as.Date("2020-01-01", "%Y-%m-%d"))), linetype="dashed") +
  labs( title = '2 Year Sales Exponential Forecast', y = 'Total Sales (Units)', x = 'Fiscal Year')
```
